---
title: "R Notebook"
output: html_notebook
---
Throughout this file, replace "yourpath" with location of files. When writing out files (these lines generally commented out), replace "filename" with your desired filename.
Setting path is sometimes not intuitive within Rmd file. May need to re-set within blocks.

```{r echo=FALSE}
rm(list = ls())
setwd("yourpath")
library("tximport")
library("limma")
library("DESeq2")
library("edgeR")

# Read in Propinquus clusters file, sample list and reads
tx2gene = readr::read_tsv("clusters.txt")
Ssamples <- read.table("Ssamples.txt", header=TRUE)
Ssamples$Group <- as.factor(Ssamples$Group)
Sfiles <- file.path(Ssamples$File, "quant.sf")
# This command is tricky. Need to run it together wit the setwd command
Limma_txi <- tximport(Sfiles, type="salmon", tx2gene=tx2gene, countsFromAbundance = "lengthScaledTPM")

y_init <- DGEList(Limma_txi$counts)

# filtering
keep <- rowSums(y_init$counts >= 15 ) >= 4
y<- y_init[keep,]
y <- calcNormFactors(y)
logcpms <- cpm(y, log=TRUE)

#For supp
cpms <- cpm(y)
#write.csv(cpms, file="filename.csv")

design <- model.matrix(~0 + Group, data = Ssamples)
v <- voomWithQualityWeights(y, design, plot=TRUE)

# For WGCNA
b <- log2(y$counts +1)
b <- t(b)
row.names(b) <- Ssamples$File
#write.csv(b, file="yourpath/CpLog2_for_WGCNA.csv")
```

```{r}
setwd("yourpath")
library("tximport")
library("DESeq2")
Stxi <- tximport(Sfiles, type="salmon", tx2gene=tx2gene)
#Create DESeqDataSet, 
SddsTxi <- DESeqDataSetFromTximport(Stxi,
                                    colData = Ssamples,
                                    design = ~ Group)
dds <- DESeq(SddsTxi)
keep <- rowSums(counts(dds)>= 10) >=4
dds <- dds[keep,]
```


```
```{r}
# Run comparisons in limma of each site against mean. 
# First make contrast matrix 

cmat <- makeContrasts( 
  S1 = (GroupS1 - (GroupS2 + GroupS3 + GroupS4)/3),
  S2 = (GroupS2 - (GroupS1 + GroupS3 + GroupS4)/3),
  S3 = (GroupS3 - (GroupS2 + GroupS1 + GroupS4)/3),
  S4 = (GroupS4 - (GroupS2 + GroupS3 + GroupS1)/3),
  levels = colnames(design))
colnames(cmat) <- abbreviate(colnames(cmat))
cmat

# Now identify, count and write out DEGs
library(tidyverse)
fit <-lmFit(v, design)
fit2 <- contrasts.fit(fit, cmat)
fit2 <- eBayes(fit2)
top1 <- as.data.frame(topTable(fit2, coef=NULL, adjust = "fdr", sort.by="B", number=30000))
#write.csv(top1, file="yourpath/SiteMeanDE_table.csv")

#Correcting for multiple comparisons
results2 <- decideTests(fit2, p.value=0.05, method="separate", adjust.method="BH")
cat("\n DE with global correction for multiple comparisons\n")
summary(results2)
results2 <- as.data.frame(results2)

Annot <- read.csv("yourpath/Supp_Calanus_Annotation_with_WGCNA.csv", header=TRUE)
wanted <- row.names(top1)
top1$Cluster <-wanted
Annot %>% filter(Cluster %in% wanted) -> my_output
my_output <- my_output[,c(2:4,6:13)]
top = merge(top1, my_output, by = "Cluster")

Site1_results <- as.data.frame(topTable(fit2, coef=1, adjust = "fdr",  number=Inf)) 
Site1_results <- cbind(rownames(Site1_results), data.frame(Site1_results, row.names=NULL))
names(Site1_results)[names(Site1_results) == 'rownames(Site1_results)'] <- 'Cluster'
Site1_results_annot <- merge(Site1_results, my_output, by='Cluster')
S1_DE <- row.names(filter(results2, `S1`==1 | `S1`==-1))
S1_DEG <- Site_results_annot[Site_results_annot$Cluster %in% S1_DE, ]
#write.csv(S1_DEG, file="yourpath/S1_DE.csv")

Site2_results <- as.data.frame(topTable(fit2, coef=2, adjust = "fdr",  number=Inf)) 
Site2_results <- cbind(rownames(Site2_results), data.frame(Site2_results, row.names=NULL))
names(Site2_results)[names(Site2_results) == 'rownames(Site2_results)'] <- 'Cluster'
Site2_results_annot <- merge(Site2_results, my_output, by='Cluster')
S2_DE <- row.names(filter(results2, `S2`==1 | `S2`==-1))
S2_DEG <- Site2_results_annot[Site2_results_annot$Cluster %in% S2_DE, ]
#write.csv(S2_DEG, file="yourpath/S2_DE.csv")

Site3_results <- as.data.frame(topTable(fit2, coef=3, adjust = "fdr",  number=Inf)) 
Site3_results <- cbind(rownames(Site3_results), data.frame(Site3_results, row.names=NULL))
names(Site3_results)[names(Site3_results) == 'rownames(Site3_results)'] <- 'Cluster'
Site3_results_annot <- merge(Site3_results, my_output, by='Cluster')
S3_DE <- row.names(filter(results2, `S3`==1 | `S3`==-1))
S3_DEG <- Site3_results_annot[Site3_results_annot$Cluster %in% S3_DE, ]
#write.csv(S3_DEG, file="yourpath/S3_DE.csv")

Site4_results <- as.data.frame(topTable(fit2, coef=4, adjust = "fdr",  number=Inf)) 
Site4_results <- cbind(rownames(Site4_results), data.frame(Site4_results, row.names=NULL))
names(Site4_results)[names(Site4_results) == 'rownames(Site4_results)'] <- 'Cluster'
Site4_results_annot <- merge(Site4_results, my_output, by='Cluster')
S4_DE <- c(row.names(filter(results2, `S4`==1 | `S4`==-1)))
S4_DEG_test <- Site4_results_annot[Site4_results_annot$Cluster %in% S4_DE, ]
S4_DEG <- Site4_results_annot[Site4_results_annot$Cluster %in% S4_DE, ]
#write.csv(S4_DEG_test, file="yourpath/S4_DE.csv")

```

```{r}
# For GoMWU
Site1_results <- as.data.frame(topTable(fit2, coef=1, adjust = "fdr",  number=Inf)) 
write.csv(Site1_results, file="yourpath/CpS1_DEGs.csv")
Site2_results <- as.data.frame(topTable(fit2, coef=2, adjust = "fdr",  number=Inf))  
write.csv(Site2_results, file="yourpath/CpS2_DEGs.csv")
Site3_results <- as.data.frame(topTable(fit2, coef=3, adjust = "fdr", number=Inf))
write.csv(Site3_results, file="yourpath/CpS3_DEGs.csv")
Site4_results <- as.data.frame(topTable(fit2, coef=4, adjust = "fdr", number=Inf))  
write.csv(Site4_results, file="yourpath/CpS4_DEGs.csv")


```

